%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%                                              %%%%%%
%%%%  Author: Peter Wilson                        %%%%%%
%%%%                                              %%%%%%
%%%%  ANDES element as programmed in Kratos                        %%%%%%
%%%%                                              %%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter[ANDES-DKQ class in Kratos]{ANDES-DKQ class in \\ Kratos}
\label{app:ANDES element as programmed in Kratos}
\renewcommand{\Thema}{ANDES-DKQ class in Kratos}

The ANDES-DKQ element is implemented in Kratos under the class \texttt{ShellThinElement3D4N} as per the established Kratos element naming convention. This appendix provides a brief description of all methods (arguments excluded for brevity) and variables within class \\ \texttt{ShellThinElement3D4N}.

\section{\texttt{ShellThinElement3D4N} methods}
\subsection{Public methods}
\texttt{ShellThinElement3D4N()}: Overloaded element constructor. Non-linear geometry bool has the default argument of \texttt{false}.

\texttt{\textasciitilde ShellThinElement3D4N()}: Element destructor.

Remaining public methods have names and overall descriptions as per \texttt{ShellThickElement3D3N}'s public methods.

\subsection{Private methods}
\texttt{CalculateStressesFromForceResultants()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateLaminaStrains()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateLaminaStresses()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateTsaiWuPlaneStress()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateVonMisesStress()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CheckGeneralizedStressOrStrainOutput()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{DecimalCorrection()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{SetupOrientationAngles()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{InitializeCalculationData()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateBMatrix()}: Calculates the strain-displacement B matrix at the current gauss point.

\texttt{CalculateSectionResponse()}: Calculates section response, including stress computation from available strains, according to the constitutive law assigned.

\texttt{CalculateGaussPointContribution()}: Calculates the current gauss point's contribution to the element stiffness matrix.

\texttt{AddBodyForces()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{CalculateAll()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{TryGetValueOnIntegrationPoints\_MaterialOrientation()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{TryGetValueOnIntegrationPoints\_GeneralizedStrainsOrStresses()}: As per identically named \texttt{ShellThickElement3D3N} method.

\texttt{printMatrix()}: As per identically named \texttt{ShellThickElement3D3N} method.

\section{\texttt{ShellThinElement3D4N} member variables}
\subsection{Public}
class \texttt{JacobianOperator}: Provides commonly used Jacobian functionality. Detailed below.

\subsection{Private}
class \texttt{CalculationData}: Contains the bulk of element data necessary for stiffness matrix computation. Detailed below.

\texttt{mpCoordinateTransformation}: Instance of the element's coordinate transformation. May be linear or non-linear corotational.

\texttt{mSections}: Container for the cross section instances associated with each integration point.

\texttt{mThisIntegrationMethod}: Current integration method of the element.

\texttt{mOrthotropicSectionRotation}: In-plane rotation angle for the element material cross section. Default is 0.0.

\subsection{Public, class \texttt{JacobianOperator} methods}
All methods described in this sub-section are scoped to \texttt{JacobianOperator} unless noted otherwise.

\texttt{JacobianOperator()}: Class constructor

\texttt{\textasciitilde JacobianOperator()}: Class destructor.

\texttt{Jacobian()}: Returns Jacobian matrix at the current gauss point.

\texttt{Inverse()}: Returns inverse Jacobian matrix at the current gauss point.

\texttt{XYDerivatives()}: Returns a matrix of Cartesian shape function derivatives at the current gauss point.

\texttt{Determinant()}: Returns the determinant of the Jacobian matrix at the current gauss point.

\subsection{Public, class \texttt{JacobianOperator} member variables}
All variables described in this sub-section are scoped to \texttt{JacobianOperator} unless noted otherwise.

\texttt{mJac}: Jacobian matrix

\texttt{mInv}: Inverse Jacobian matrix

\texttt{mXYDeriv}: Matrix of Cartesian shape function derivatives.

\texttt{mDet}: Determinant of the Jacobian matrix.

\subsection{Private, class \texttt{CalculationData} member variables}
All variables described in this sub-section are scoped to \texttt{CalculationData} unless noted otherwise.

\texttt{LCS0}: Reference coordinate system of undeformed configuration.

\texttt{LCS}: Current coordinate system of deformed configuration.

\texttt{s\_xi}: $\xi$ unit vector in Cartesian coordinates.

\texttt{s\_eta}: $\eta$ unit vector in Cartesian coordinates.

\texttt{r\_cartesian}: Array of Cartesian point positions.

\texttt{dA}: Array of integration areas associated with each gauss point. Pre-multiplied with integration weight, equal to $det(\mathbf{J}_i) w_i$.

\texttt{globalDisplacements}: Global displacement vector.

\texttt{localDisplacements}: Local displacement vector.

\texttt{CalculateRHS}: Bool to calculate internal force vector.

\texttt{CalculateLHS}: Bool to calculate stiffness matrix.

\texttt{basicQuad}: Bool to use a basic displacement-derived membrane formulation as per equation \ref{eqApp1_5}. Default value is \texttt{false}.

\texttt{N}: Shape function vector at the current integration point.

\texttt{gpIndex}: Integer holding the index of the current gauss point.

\textit{ANDES membrane formulation data}

\texttt{alpha}: Constant parameter alpha in the ANDES basic membrane formulation as per equation \ref{equation07}. Value is 1.5.

\texttt{L\_mem}: ANDES basic membrane lumping matrix. Refer equation \ref{equation06}.

\texttt{H\_mem\_mod}: ANDES higher order filter matrix as per Haugen's DOF ordering. Refer equation \ref{equation11}.

\texttt{Z}: Transformation matrix between Haugen's and Kratos' DOF ordering for the higher order filter matrix as per equation \ref{equation18_2}.

\texttt{B\_h\_1}: Higher order membrane B-matrix of node 1, as per equation \ref{equation17}.

\texttt{B\_h\_2}: Higher order membrane B-matrix of node 2, as per equation \ref{equation18}.

\texttt{B\_h\_3}: Higher order membrane B-matrix of node 3, as per equation \ref{equation17}.

\texttt{B\_h\_4}: Higher order membrane B-matrix of node 4, as per equation \ref{equation18}.

\texttt{B\_h\_bar}: Mean higher order membrane B-matrix, as per equation \ref{eqnAndesDevB}.

\texttt{T\_13}: Transformation matrix as per equation \ref{equation17}.

\texttt{T\_24}: Transformation matrix as per equation \ref{equation18}.

\textit{DKQ bending formulation data}

\texttt{DKQ\_a}: Vector of DKQ formulation coefficients $a_k$ as per equation \ref{equation23}.

\texttt{DKQ\_b}: Vector of DKQ formulation coefficients $b_k$ as per equation \ref{equation23}.

\texttt{DKQ\_c}: Vector of DKQ formulation coefficients $c_k$ as per equation \ref{equation23}.

\texttt{DKQ\_d}: Vector of DKQ formulation coefficients $d_k$ as per equation \ref{equation23}.

\texttt{DKQ\_e}: Vector of DKQ formulation coefficients $e_k$ as per equation \ref{equation23}.

\texttt{DKQ\_indices}: Matrix of DKQ mapping indices $i,\ r,\ s$ ordering as per commentary beneath equation \ref{equation26}.

\texttt{DKQ\_invJac}: Vector of inverse Jacobian matrices for each gauss point as per equation \ref{equation29}.

\texttt{DKQ\_jac}: Vector of Jacobian matrices for each gauss point as per equation \ref{equation29}.

\texttt{DKQ\_jac\_det}: Vector of Jacobian determinants for each gauss point as per equation \ref{equation29}.

\textit{General calculation-variable data}

\texttt{B}: Element strain-displacement B matrix.

\texttt{D}: Element material matrix.	

\texttt{BTD}: Auxiliary matrix to store the product $\mathbf{B}^T \mathbf{D}$.	

\texttt{generalizedStrains}: Element strain vector.

\texttt{generalizedStresses}: Element stress vector.

\texttt{jacOp}: Instance of \texttt{JacobianOperator} class.

\texttt{rlaminateStrains}: Vector of element lamina strain vectors.

\texttt{rlaminateStresses}: Vector of element lamina stress vectors.

\texttt{SectionParameters}: Parameters of the shell cross section calculation.

\texttt{CurrentProcessInfo}: Reference to the Kratos job's current process information.